---
title: "Homework6"
author: "Yen Wang and Nathan Frank"
date: "2/28/2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)

data(happy, package="productplots")
head(happy)
HAPPY <- readRDS("data/HAPPY.rds")
HAPPY$HAPPY = as.factor(HAPPY$HAPPY)
```

Data cleaning: the values “IAP”, “DK” and “NA” all encode missing values. 
Replace all of these instances by the value NA.
```{r}
HAPPY = replace(HAPPY, HAPPY == "IAP", NA)
HAPPY = replace(HAPPY, HAPPY == "DK", NA)
HAPPY = replace(HAPPY, HAPPY == "NA", NA)
```

Check the type of the variable and cast into the right type (factor variable for categorical variables).
For age, change “89 OR OLDER” to 89 and assume the variable should be numeric.
Bring all levels of factors into a sensible order. For marital you could e.g. order the levels according to average age.

```{r}
str(HAPPY)
happy1 = 
 HAPPY %>% mutate(AGE = replace(AGE, AGE == "89 OR OLDER", 89)) %>% 
  mutate(AGE = as.numeric(AGE), 
        HAPPY = as.factor(HAPPY), 
        SEX = as.factor(SEX), 
        MARITAL = as.factor(MARITAL),
        FINRELA = as.factor(FINRELA),
        HEALTH = as.factor(HEALTH),
        PARTYID = as.factor(PARTYID),
        POLVIEWS = as.factor(POLVIEWS),
        DRGREE = as.factor(DEGREE)) %>% 
          mutate(DEGREE = factor(DEGREE, levels = c("LT HIGH SCHOOL", "HIGH SCHOOL", "JUNIOR COLLEGE", "BACHELOR", "GRADUATE")))


levels(happy1$MARITAL) = c("never married", "married", "separated", "divorced", "widowed")
levels(happy1$FINRELA) = c("far below average","below average","average","above average","far above average")
levels(happy1$HEALTH) = c("poor","fair","good","excellent")
levels(happy1$PARTYID) = c("strong republican", "not str republican", "ind,near rep", "independent", "ind,near dem", "not str democrat", "strong democrat", "other party")
levels(happy1$POLVIEWS) = c("extrmly conservative", "conservative","slghtly conservative","moderate","slightly liberal","liberal","extremely liberal")
#levels(happy1$DEGREE) = c("LT HIGH SCHOOL", "HIGH SCHOOL", "JUNIOR COLLEGE", "BACHELOR", "GRADUATE")

str(happy1)
levels(happy1$HEALTH)

```



```{r}
#happiness by degree
happy.degree = happy1 %>% 
  group_by(DEGREE) %>% summarise(m.degree = mean(as.numeric(HAPPY), na.rm = TRUE))

happy.degree %>% ggplot(aes(x = DEGREE, y = m.degree)) + geom_point()
```


```{r}
#happiness by finacial relationship
happy.finrela = happy1 %>% 
  group_by(FINRELA) %>% summarise(m.finrela = mean(as.numeric(HAPPY), na.rm = TRUE))

happy.finrela %>% ggplot(aes(x = FINRELA)) + geom_bar(aes(weight = m.finrela), position="dodge") +
  labs(title="Happiness by Financial Relationship", y="Happiness")
```